---
title: "covid 19"
author: "Melany Vasquez"
date: "summer 2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata$sex <- as.factor(kRdata$sex)
kRdata$result <- as.factor(kRdata$result)


```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent, use ggplot commands to do this

```{r}
dd2 <- kRdata %>% group_by(sex,result) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(dd2,aes(x=result,y=count,fill=sex))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(dd2,aes(x=result,y=prcnt*100,fill=sex)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

Theses current bargraphs about 100% of women are alive after covid, men arent too far behind 100% either. However results show a slighter bigger amount of dead males rather than women.

```{r}
barchartGC(~result + Sex,data=kRdata)
barchartGC(~result + Sex,data=kRdata, type="percent")

```

The barchart represent that male deaths related to covid make up about 60% of the total death and females make up the other 40% of deaths.But it shows a fair amount of women who also die from covid.
  
## Numerical results

```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Our numerical results shows us that out of 100, males make up 64% of covid related death compared to females at 36%. This also shows that males are twice as likely to die from covid compared to females.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

Fishers exact test for count data shows that males are 2.3 times more likely to die from covid compared to women. Our p-value from the fishers exact test is 0.0009 which is a very small number this allows me to rejecet the null hypothesis that states covid deaths are independent of sex.


## Conclusions

The null hypothesis stated that covid related deaths were independent to sex. In order to figure out for starters we used our Cat_Cat helper file for the graphical purposes in order to determine the variables : results, and sex.First we added a barchart in order to show where the amount of deaths per 100. From here it shows that roughly about 60% of deaths out of 100 were males and for females it was 40%. Then we ran the numerical results in order to have a precise number of covid related death between male and female. This gave us that 64 males die to covid compared to 36 females. This also means that we can state that males were more likely to die at a higher rate than females when catching covid-19. 
